---
title: "PeerAssessment"
output: html_document
---

First load the data and change the date into the correct format 

```{r, echo=TRUE}
activity = read.csv("activity.csv")
activity$date = as.Date(activity$date, format = "%m/%d/%Y")
```


Calculate the total number of steps taken per day 
```{r, echo=TRUE}
total = aggregate(activity$steps ~ as.factor(activity$date), FUN=sum)
colnames(total) = c("date", "totalSteps")
hist(total$totalSteps, breaks = 15, col = "lightblue", xlab = "Total steps per day", main = "Histogram of total steps per day")

```

The mean total number of steps taken is 
```{r, echo=TRUE}
mean(total$totalSteps)
```

The median total number of steps taken is 
```{r, echo=TRUE}
median(total$totalSteps)
```

Time series plot of 5-min interval and the average number of steps taken

```{r, echo=TRUE}
yvals = aggregate(activity$steps ~ activity$interval, FUN=mean)
colnames(yvals) = c("interval", "totalSteps")

plot(yvals$interval, yvals$totalSteps, type="l", xlab = "Interval", ylab = "Average total steps per day", main = "Average steps per interval")
```

5-minute interval with maximum number of steps
```{r, echo=TRUE}
yvals$interval[which(yvals$totalSteps == max(yvals$totalSteps))]
```

Finding total number of NAs in the data set 
```{r, echo=TRUE}
sum(is.na(activity$steps))
```

Replace NAs with mean in each interval 
```{r, echo=TRUE}

```

New Factor variable with weekday/weekend
```{r, echo=TRUE}
weekends = as.factor(weekdays(activity$date) %in% c('Sunday', 'Saturday'))
activity$weekends = weekends 
par(mfrow=c(2,1))
week = activity[which(activity$weekends==F),]
weekdays.val = aggregate(week$steps ~ week$interval, FUN=mean)
colnames(weekdays.val) = c("interval", "steps")
end = activity[which(activity$weekends==T),]
weekends.val = aggregate(end$steps ~ end$interval, FUN = mean)
colnames(weekends.val) = c("interval", "steps")
```

Plotting Weekend vs weekday steps 
```{r, echo=TRUE}
plot(weekdays.val$steps ~ weekdays.val$interval, type = "l", xlab="Interval", main = "Weekdays", ylab = "Steps")
plot(weekends.val$steps ~ weekends.val$interval, type = "l", xlab="Interval", main = "Weekends", ylab = "Steps")
```
